Advantages of mixed unitary operators for quantum information processing

### 2018-12-28 decorrelation/mixed_unitaries.tex
more structural changes

The \emph{size} of an error in quantum gates may  be quantified in a number of ways. Two of the most common metrics are the average gate fidelity, $\mathcal{F}$, and the diamond norm, $\vert\vert\cdot\vert\vert_\diamond$. These may be represented in terms of the error maps as
\begin{equation}
	\vert\vert \error \vert\vert_\diamond = \sup_\rho \vert \vert (I\otimes I)(\rho) - (\error \otimes I)(\rho) \vert\vert_1
\end{equation}
\begin{equation}
	\mathcal{F}(\error) = \frac{Tr(\error^*\otimes\error) + d}{d^2 + d}
\end{equation}
The diamond norm is generally linear in the over-rotation angle of a quantum operation, while the AGI is generally quadratic in the over-rotation angle of a quantum operation.

@@

\end{center}

@@

In this case, the MUP that minimizes the diamond norm is given by:
\begin{equation}
\frac{1}{2}U_{+\epsilon}^*\otimes U_{+\epsilon} + \frac{1}{2}U_{-\epsilon}^*\otimes U_{-\epsilon}
\end{equation}
Additionally, it is clear that because the AGI is linear, the AGI of any convex weighting of these two channels will be $\epsilon^2$.
\note{This example is really useful and should be expanded. We should discuss the various optimization problems, a mixed unitary strategy, the fidelity of the resulting channel.}

% ==============================================================================
% Section: Robustly Mixed Unitary Processes
% ==============================================================================
\section{Robustly Mixed Unitary Processes}
\label{sec:robustly_mixed}
\note{Decide on naming convention, Anthony has used $\ell$ MUP in the rest of the text}

\note{Anthony references this section when discussing minimizing the offdiagonal elements of the error's process matrix.}

@@

Following from Lemma 2 in \cite{Campbell2017}, if $0\in $ Conv$[\{\tilde{H}_i(t, \vec{\delta})\}]$, where Conv is the convex hull of its arguments, then we know that $\vec{\omega}$ exists that quadratically decreases the diamond norm. Then, following from the mixing lemma proven by both Campbell and Hastings\cite{1612.01011}, we can bound the the diamond distance from a balanced control to $U_T$ by $\epsilon^2$. We can relax Equation \ref{eq:MUP}, to be
\begin{equation}\label{eq:RBC-relaxed}
\begin{gathered}
\sum_k\omega_kD^n_k = 0\\
\end{gathered}
\end{equation}
In this context, since we are assuming that the derivatives are all of norm $\epsilon$, the above condition guarantees that the error unitary be no worse than $\sum^n\epsilon^2\delta^n$ near the origin -- all orders can be quadratically supressed. Figure \ref{fig:vectorspace} gives geometric intuition for robustly mixed unitary processes.

@@

% ==============================================================================
% Section: Generationg of Robustly Mixed Unitary Processes
% ==============================================================================
\section{Generation of Robustly Mixed Unitary Processes}
\label{sec:generation}

\subsection{Random Gate Synthesis}
\label{sec:random_gate_synthesis}
 In this article, we use optimal control to generate the one qubit families of controls for MUPs. One could use any of the many available quantum optimal control techniques \cite{Khaneja2005, Caneva2011, Machnes2018}, and for our numerics we chose to use the GRAPE algorithm. First described in \cite{Khaneja2005}, the GRAPE (GRadient Ascent Pulse Engineering) algorithm is a technique for finding piecewise constant control sequences that approximate a desired unitary, $U_T$, given a Hamiltonian with controlled and uncontrolled terms. Defining the uncontrolled Hamiltonian as $H_0$, the control Hamiltonians as $H_{i\neq 0}$, and the \textit{control matrix} $c_{ij}$ as containing the control amplitude associated with the $i^{th}$ time step and the $j^{th}$ Hamiltonian, we can write an the unitary for any timestep of evolution as:
\begin{equation}\label{eq:3}
  U_i = \exp\{-i\Delta t(H_0 + \sum_{j=1}^{n}c_{ij}H_{j}\}
\end{equation}
To measure the simularity of our approxiate unitary $U=\prod_iU_{i=1}^n$ to our target unitary $U_T$, we can define a cost function $J(U) = Tr\{U_T^{\dagger}U\}$.

@@

\subsection{Optimal Weighting}
To generate robustly mixed unitary processes, we can solve the following convex optimization problem:

\begin{equation}\label{eq:minimization}
  \begin{split}
    &\underset{\omega_j\geq0, |\omega|_1=1}{\textbf{minimize}: } ||D_{\ell}^T\omega||\\
    &\textbf{subject to: } \forall i<\ell, \sum \omega_jD_j^i=0\\
  \end{split}
\end{equation}

If the control family is too small a solution won't exist, in particular as in \cite{Campbell2017} we need  $0\in $ Conv$[D_j^i]$.  This can be accomplished by modifying the oracle in Campbell's algorithm to produce robust controls, in particular by updating the GRAPE cost function to include derivative information, one could produce controls with derivatives of order $\epsilon$. In our case, by using the Gaussian weighted cost function, we ensure that the controls generated should have vanishing first derivative.

This minimization problem, while sufficient for quadratically decreasing the diamond norm relative to the \textit{worst} controls in the family, does not preferentially select the controls with the least error. That is to say, as in Section \ref{sec:mixed_unitary_processes}, both \{$U_{+2\epsilon}$, $U_{-2\epsilon}$\} and \{$U_{\epsilon}$, $U_{-\epsilon}$\} minimize this cost function for a 0MUP. To encourage the inclusion of controls with smaller error, we may impose an $\ell_2$ penalty to our cost function by rewriting it as:

\begin{equation}\label{eq:minimization_l2}
  \begin{split}
    &\underset{\omega_j\geq0, |\omega|_1=1}{\textbf{minimize}: } ||D_{\ell}^T\omega||_2 + \eta\sum||D_0^T\omega||_2\\
  \end{split}
\end{equation}
with $\eta \geq 0$.

To make sure the routine remains practical, we would also like to regularize our objective function to enforce sparsity. In this case, lasso regularization \cite{tibshirani1996regression} is insufficient as we already constrain the one norm of the vector we optimize over to be one. However, the problem of enforcing sparsity in such situations has been considered in \cite{NIPS2012_4504} and can be expressed via another convex program that extends Equation \ref{eq:minimization}:
\begin{equation}
\begin{split}
\underset{n\in[N]}{\textbf{minimize}}\{
    &\underset{\omega_j\geq0, |\omega|_1=1, t\geq0}{\textbf{minimize}: } ||D_{\ell}^T\omega|| + t\\
    &\textbf{subject to: } \omega_n > \frac{\lambda}{t}\\
    &\phantom{\textbf{subject to: }} \forall i<\ell, \sum \omega_jD_j^i=0\}
\end{split}
\end{equation} where $\lambda$ is a hyperparameter to be optimized over.

% ==============================================================================
% Section: Numerical Results
% ==============================================================================
\section{Numerical Results}
\label{sec:numerical_results}
In the following numerical results, we explore using the methods in Section \ref{sec:mixed_unitary_processes} to build MUPs. We consider the following model for a single tunable qubit: 
\begin{equation}\label{eq:1Qham}
  H(\delta, \epsilon, t) = \epsilon\sigma_z + (1 + \delta)(c_x(t)\sigma_x + c_y(t)\sigma_y)
\end{equation}
We use the GRAPE algorithm as discussed in Section  \ref{sec:random_gate_synthesis} with N=25 steps and total evolution time of $\pi$ to generate 100 candidate controls, with a standard deviation of $\sigma=.001$ for the distribution in Equation \ref{eq:quadrature}. We assume that the errors on $\sigma_x$ and $\sigma_y$ are perfecly correlated, as is the case in systems that implement RZ rotations with phase shifts of the control signal. Solving the optimization problem defined in Section \ref{sec:robustly_mixed} yields similar MUPs for $RX(\frac{\pi}{2})$ and $RY(\frac{\pi}{2})$, with the results for $RY(\frac{\pi}{2})$ shown in Figure \ref{fig:YMUP}. These results demonstrate several properties that make MUPs both useful and tractable.

@@

First, naively generating a 0MUP results in nontrivial support on all the members of the control family. However, by rewriting the minimization to impose the sparsity constraint discussed in Section \ref{sec:robustly_mixed}, we can generate a 0MUP that uses just five of the controls. This shows that through adding constraints to our optimization routine, we can make the MUP practically useful. In both cases we impose the same $\ell_2$ penalty as described in Section \ref{sec:robustly_mixed}, so that the algorithm preferentially selects controls with smaller errors. Imposing this constraint allows us to trade off flatness at the origin for performance.

In our two-qubit example we consider the following model for two tunable qubits coupled by a resonant exchange interaction, similar to that in \cite{McKay2016}:
\begin{equation} \label{eq:2Qham}
\begin{split}
H(\vec{\delta}, \vec{\epsilon}, t) = &\sum_{j=1}^2(\epsilon_j\sigma_z^j + (1 + \delta_j)(c_x^jx(t)\sigma_x^j + c_y^j(t)\sigma_y^j)) \\
&+ \frac{1}{10}(XX + YY)
\end{split}
\end{equation}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\columnwidth]{rb_data.pdf}
  \caption{Randomized benchmarking experiments ran using different pulse definitions. The four plots on the left are from the incorrectly calibrated pulse, while the top right is the calibrated pulse, and the bottom right is the BCS.}
  \label{fig:rb}
\end{figure}

>>>>

The \emph{size} of an error in quantum gates may  be quantified in a number of ways. Two of the most common metrics are the average gate fidelity, $\mathcal{F}$, and the diamond norm, $\vert\vert\cdot\vert\vert_\diamond$. These may be represented in terms of the error maps as
\begin{equation}
	\vert\vert \error \vert\vert_\diamond = \sup_\rho \vert \vert (I\otimes I)(\rho) - (\error \otimes I)(\rho) \vert\vert_1
\end{equation}
\begin{equation}
	\mathcal{F}(\error) = \frac{d^2 - \tr{(\error^*\otimes\error)}}{d^2 + d}
\end{equation}
The diamond norm is generally linear in the over-rotation angle of a quantum operation, while the AGI is generally quadratic in the over-rotation angle of a quantum operation.

@@

\end{center}
\note{This example is really useful and should be expanded. We should discuss the various optimization problems, a mixed unitary strategy, the fidelity of the resulting channel.}

@@

In this case, the MUP that minimizes the diamond norm is given by:
\begin{equation}
\frac{1}{2}U_{+\epsilon}^*\otimes U_{+\epsilon} + \frac{1}{2}U_{-\epsilon}^*\otimes U_{-\epsilon}
\end{equation}
Additionally, it is clear that because the AGI is linear, the AGI of any convex weighting of these two channels will be $\epsilon^2$. More generally, the problems remains: given a collection of channels, how can one efficiently compute a weighting that minimizes a particular metric.

As discussed in \cite{Campbell2017}, a sufficient condition is:
\begin{equation}\label{eq:campbell-condition}
\sum \omega_j H_j = 0
\end{equation}

This condition can be satisfied by solving the following convex optimization problem:

\begin{equation}\label{eq:minimization}
  \begin{split}
    &\underset{\omega_j\geq0, |\omega|_1=1}{\textbf{minimize}: } ||H^T\omega||_2\\
  \end{split}
\end{equation}

where $H$ is the matrix whose rows are the vectorized Hamiltonians.

This minimization problem, while sufficient for quadratically decreasing the diamond norm relative to the \textit{worst} controls in the collection, does not preferentially select the controls with the least error. That is to say, both \{$U_{+2\epsilon}$, $U_{-2\epsilon}$\} and \{$U_{\epsilon}$, $U_{-\epsilon}$\} minimize this Equation\ref{eq:minimization}. To encourage the inclusion of controls with smaller error, we may impose an $\ell_2$ penalty to our cost function by rewriting it as:

\begin{equation}\label{eq:minimization_l2}
  \begin{split}
    &\underset{\omega_j\geq0, |\omega|_1=1}{\textbf{minimize}:  }||H^T\omega||_2 + \eta\sum||H||_2^T\omega\\
  \end{split}
\end{equation}
with $\eta \geq 0$.

%\begin{equation}\label{eq:minimization_l2}
%  \begin{split}
%    &\underset{\omega_j\geq0, |\omega|_1=1}{\textbf{minimize}: } ||D_{\ell}^T\omega||_2 + \eta\sum||D_0^T\omega||_2\\
%  \end{split}
%\end{equation}

To make sure the routine remains practical, we would also like to regularize our objective function to enforce sparsity. In this case, lasso regularization \cite{tibshirani1996regression} is insufficient as we already constrain the one norm of the vector we optimize over to be one. However, the problem of enforcing sparsity in such situations has been considered in \cite{NIPS2012_4504} and can be expressed via another convex program that extends Equation \ref{eq:minimization}:

\begin{equation}\label{eq:minimization_regularization}
\begin{split}
\underset{n\in[N]}{\textbf{minimize}}\{
    &\underset{\omega_j\geq0, |\omega|_1=1, t\geq0}{\textbf{minimize}: } ||H^T\omega|| + t\\
    &\textbf{subject to: } \omega_n > \frac{\lambda}{t}\}
\end{split}
\end{equation} where $\lambda$ is a hyperparameter to be optimized over.

%\begin{equation}
%\begin{split}
%\underset{n\in[N]}{\textbf{minimize}}\{
%   &\underset{\omega_j\geq0, |\omega|_1=1, t\geq0}%{\textbf{minimize}: } ||D_{\ell}^T\omega|| + t\\%    &\textbf{subject to: } \omega_n > \frac{\lambda}{t}\\
%    &\phantom{\textbf{subject to: }} \forall i<\ell, \sum \omega_jD_j^i=0\}
%\end{split}
%\end{equation} where $\lambda$ is a hyperparameter to be optimized over.

% ==============================================================================
% Section: Robustly Mixed Unitary Processes
% ==============================================================================
\section{Robustly Mixed Unitary Processes}
\label{sec:robustly_mixed}

@@

Following from Lemma 2 in \cite{Campbell2017}, if $0\in $ Conv$[\{\tilde{H}_i(t, \vec{\delta})\}]$, where Conv is the convex hull of its arguments, then we know that $\vec{\omega}$ exists that quadratically decreases the diamond norm. Then, following from the mixing lemma proven by both Campbell and Hastings\cite{1612.01011}, we can bound the the diamond distance from a balanced control to $U_T$ by $\epsilon^2$. We can relax Equation \ref{eq:MUP}, to be
\begin{equation}\label{eq:MUP-relaxed}
\begin{gathered}
\sum_k\omega_kD^n_k = 0\\
\end{gathered}
\end{equation}
In this context, since we are assuming that the derivatives are all of norm $\epsilon$, the above condition guarantees that the error unitary be no worse than $\sum^n\epsilon^2\delta^n$ near the origin -- all orders can be quadratically supressed. Figure \ref{fig:vectorspace} gives geometric intuition for the conditions required to produce an $\ell MUP$.

@@

To generate robustly mixed unitary processes, we can solve the following convex optimization problem, generalizing Equation \ref{eq:minimization}:

\begin{equation}\label{eq:robust_minimization}
  \begin{split}
    &\underset{\omega_j\geq0, |\omega|_1=1}{\textbf{minimize}: } ||D_{\ell}^T\omega||\\
    &\textbf{subject to: } \forall i<\ell, \sum \omega_jD_j^i=0\\
  \end{split}
\end{equation}

Additionally, this can be extended in the obvious way to use Equations \ref{eq:minimization_l2} and\ref{eq:minimization_regularization}.

% ==============================================================================
% Section: Numerical Results
% ==============================================================================
\section{Numerical Results}
\label{sec:numerical_results}
\subsection{Random Gate Synthesis}
\label{sec:random_gate_synthesis}
 In this article, we use optimal control to generate the one qubit families of controls for MUPs. One could use any of the many available quantum optimal control techniques \cite{Khaneja2005, Caneva2011, Machnes2018}, and for our numerics we chose to use the GRAPE algorithm. First described in \cite{Khaneja2005}, the GRAPE (GRadient Ascent Pulse Engineering) algorithm is a technique for finding piecewise constant control sequences that approximate a desired unitary, $U_T$, given a Hamiltonian with controlled and uncontrolled terms. Defining the uncontrolled Hamiltonian as $H_0$, the control Hamiltonians as $H_{i\neq 0}$, and the \textit{control matrix} $c_{ij}$ as containing the control amplitude associated with the $i^{th}$ time step and the $j^{th}$ Hamiltonian, we can write an the unitary for any timestep of evolution as:
\begin{equation}\label{eq:3}
  U_i = \exp\{-i\Delta t(H_0 + \sum_{j=1}^{n}c_{ij}H_{j}\}
\end{equation}
To measure the simularity of our approxiate unitary $U=\prod_iU_{i=1}^n$ to our target unitary $U_T$, we can define a cost function $J(U) = Tr\{U_T^{\dagger}U\}$.

@@

If the collection of controls generated this way is too small, a solution won't exist, in particular as in \cite{Campbell2017} we need  $0\in $ Conv$[D_j^i]$.  This can be accomplished by modifying the oracle in Campbell's algorithm to produce robust controls, in this case by updating the GRAPE cost function to include derivative information, one could produce controls with derivatives of order $\epsilon$. In our case, by using the Gaussian weighted cost function, we ensure that the controls generated should have vanishing first derivative.

\subsection{One Qubit Performance Analysis}
\label{one_qubit_performance}
In the following numerical results, we explore using the methods in Section \ref{sec:mixed_unitary_processes} to build MUPs. We consider the following model for a single tunable qubit: 
\begin{equation}\label{eq:1Qham}
  H(\delta, \epsilon, t) = \epsilon\sigma_z + (1 + \delta)(c_x(t)\sigma_x + c_y(t)\sigma_y)
\end{equation}
We use the GRAPE algorithm as discussed in Section  \ref{sec:random_gate_synthesis} with N=25 steps and total evolution time of $\pi$ to generate 100 candidate controls, with a standard deviation of $\sigma=.001$ for the distribution in Equation \ref{eq:quadrature}. We assume that the errors on $\sigma_x$ and $\sigma_y$ are perfecly correlated, as is the case in systems that implement RZ rotations with phase shifts of the control signal. Solving the optimization problem defined in Section \ref{sec:robustly_mixed} yields similar MUPs for $RX(\frac{\pi}{2})$ and $RY(\frac{\pi}{2})$, with the results for $RY(\frac{\pi}{2})$ shown in Figure \ref{fig:YMUP}. These results demonstrate several properties that make MUPs both useful and tractable.

@@

First, naively generating a 0MUP results in nontrivial support on all the members of the control family. However, by rewriting the minimization to impose the sparsity constraint discussed in Section \ref{sec:robustly_mixed}, we can generate a 0MUP that uses just five of the controls. This shows that through adding constraints to our optimization routine, we can make the MUP practically useful. In both cases we impose the same $\ell_2$ penalty as described in Section \ref{sec:robustly_mixed}, so that the algorithm preferentially selects controls with smaller errors. Imposing this constraint allows us to trade off flatness at the origin for performance.
\subsection{Two Qubit Performance Analysis}
\label{two_qubit_performance}
In our two-qubit example we consider the following model for two tunable qubits coupled by a resonant exchange interaction, similar to that in \cite{McKay2016}:
\begin{equation} \label{eq:2Qham}
\begin{split}
H(\vec{\delta}, \vec{\epsilon}, t) = &\sum_{j=1}^2(\epsilon_j\sigma_z^j + (1 + \delta_j)(c_x^jx(t)\sigma_x^j + c_y^j(t)\sigma_y^j)) \\
&+ \frac{1}{10}(XX + YY)
\end{split}
\end{equation}

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{rb_data.pdf}
  \caption{Randomized benchmarking experiments ran using different pulse definitions. The four plots on the left are from the incorrectly calibrated pulse, while the top right is the calibrated pulse, and the bottom right is the BCS.}
  \label{fig:rb}
\end{figure}
